\section{Data Landscape}

\para{The Value of Data}
The newly minted role of data scientist is evidence to rise of data influence.
It used to be the case that big data was daunting. The value of data has
increased as machine learning makes big data more amenable. This increase in data value
is compounded by the democritization of cloud technologies. Microsoft, Amazon, and Google
all offer machine learning suites in cloud.


% A paragraph on how data drives business decisions, but not in the
% traditional sense. How we deal with data drives the business decisions we
% makes, eg data fedderating
\para{Data Driven Decisions}

\para{The Value of Data}

% Might not belong here?
% Need to mention that more data will make these algorithms better
% Possibly talk about the rise of imageNet and its role in computer vision
\para{Deep Learning}
Within the field of artificial intelligence, there a subfield Machine Learning.
Broadly, there are 3 kinds of machine learning: supervised, unsupervised, and
reinforcement learning. For supervised learning, our data has independent
variables and outcome variables and we train our model to learn a function that
predicts the desired outcomes. For unsupervised learning, we don't know the
outcomes that we want, but we have independent variables that might have
features that will illuminate those outcomes. Finally, in reinforcement
learning, our model is trained in an environment where it gets feedback on the
outcome it decides. When the outcome is good, the model is rewarded.

Deep learning can be supervised, unsupervised, or semi-supervised.
